---
title: "scRNA-seq_analysis_96Rik+Matn4-Combo"
output: html_notebook
---

Goal: Upload the scRNA-seq datasets generated from lncOL1 mice

```{r init, echo=FALSE}
setwd("F:/orion_copies/kallisto_out_lncol1_matn4_sort")

```

```{r}
#install.packages('knitr')
library(knitr)
#install.packages('monocle3')
library(monocle3)
#install.packages('BiocManager')
#BiocManager::install("BUSpaRse")
library(BUSpaRse)
#BiocManager::install('zealot')
library(zeallot)
#install.packages('tidyverse')
library(tidyverse)
#install.packages('DropletUtils')
#library(DropletUtils)
#install.packages('org.Mm.eg.db')
#library(org.Mm.eg.db)
library(dplyr)
#BiocManager::install("biomaRt")
library(biomaRt)
library(DropletUtils)

```
# Create monocle sc object from kallisto pipeline output
# Generate knee plots and filter each sample by a minimum UMI count per cell
```{r}
# Read Transcript->Gene mapping
tr2g <- readr::read_tsv("required_files/tr2g.tsv", col_names = FALSE)
colnames(tr2g)<-c("transcript","gene")

metadata <- read.csv("2026_LncOL1_metadata_matn4.csv")

#These sample names should align with the directory names for kallisto output
samples_ak<-metadata$sample_id[metadata$batch == 7]
samples_dh <- metadata$sample_id[metadata$batch < 7]

cds_list<-list()
# Read genes.mtx

for (i in 1:length(samples_ak)){
    print(samples_ak[i])
    mat<-t(Matrix::readMM(paste0(samples_ak[i],"/spliced/s.mtx")))
    gene_info<-readr::read_tsv(paste0(samples_ak[i],"/spliced/s.genes.txt"), col_names = FALSE)
    barcodes<-readr::read_tsv(paste0(samples_ak[i],"/spliced/s.barcodes.txt"), col_names = FALSE)
    
    cds<-new_cell_data_set(mat,
                         cell_metadata = barcodes,
                         gene_metadata = gene_info)
    
    #cds <- cds[,Matrix::colSums(exprs(cds)) != 0] 
    #cds <- estimate_size_factors(cds)
    pData(cds)$sample_id<-samples_ak[i]
    
    cds_list<-c(cds_list,cds)
}

for (i in 1:length(samples_dh)){
    print(samples_dh[i])
    mat<-t(Matrix::readMM(paste0("spliced/",samples_dh[i],"/s.mtx")))
    gene_info<-readr::read_tsv(paste0("spliced/",samples_dh[i],"/s.genes.txt"), col_names = FALSE)
    barcodes<-readr::read_tsv(paste0("spliced/",samples_dh[i],"/s.barcodes.txt"), col_names = FALSE)
    
    cds<-new_cell_data_set(mat,
                         cell_metadata = barcodes,
                         gene_metadata = gene_info)
    
    #cds <- cds[,Matrix::colSums(exprs(cds)) != 0] 
    #cds <- estimate_size_factors(cds)
    pData(cds)$sample_id<-samples_dh[i]
    
    cds_list<-c(cds_list,cds)
}

#pData(cds_list)
#fData(cds_list)
```


```{r, knee plots}
samples<-metadata$sample_id
numExpected<-4000 #cells targeted
UMI_cutoff<-1000 #expected UMI in droplets with a cell
pdf("./output/Knee-plots-test.pdf",width=6,height=6)
lapply(1:length(samples), function(sampleIdx){
    print(samples[sampleIdx])
    #tot_counts<-Matrix::colSums(exprs(cds_list[[sampleIdx]]))
    bc_rank <- barcodeRanks(exprs(cds_list[[sampleIdx]]), lower = 80)
    
    print(qplot(bc_rank$total, bc_rank$rank, geom = "line") +
      geom_vline(xintercept = metadata(bc_rank)$knee, color = "blue", linetype = 2) +
      geom_vline(xintercept = metadata(bc_rank)$inflection, color = "green", linetype = 2) +
      geom_vline(xintercept = UMI_cutoff, color = "darkred", linetype = 2) +
      geom_hline(yintercept = numExpected, color = "black", linetype = 2) +
      annotate("text", y = c(1000,5000,10000), x = 1.5 * c(metadata(bc_rank)$knee, metadata(bc_rank)$inflection,UMI_cutoff),
               label = c(
                   paste0("knee= ",metadata(bc_rank)$knee," UMIs"), 
                   paste0("inflection= ",metadata(bc_rank)$inflection," UMIs"),
                   paste0("UMI_cutoff = ",UMI_cutoff)
                   ),
                   color = c("blue", "green","darkred")) +
      scale_x_log10() +
      scale_y_log10() +
      labs(y = "Barcode rank", x = "Total UMI count") + 
      ggtitle(paste0("Knee Plot - ",samples[sampleIdx]))
    )
})
dev.off()
```

At this stage each sample is represented by one of the cds in cds_list. 
Do the filtering at specified cutoffs for each of the cds object, and then merge into one cell data set object.
```{r}
# Filter emtpy droplets, this parameter is set based on knee plots / expected cell numbers
UMI_cutoff<-500 
#Note-Started with 500, but increased this cutoff later to 1000 based on the appearance of the knee plots
filtered_cds_list<-lapply(cds_list,function(cds){
  pData(cds)$Total_UMIs<-Matrix::colSums(exprs(cds))
  #bc_rank <- barcodeRanks(exprs(cds))
  #print(metadata(bc_rank)$knee)
  #print(metadata(bc_rank)$inflection) 
  filtered_cds<-cds[, pData(cds)$Total_UMIs > UMI_cutoff]
  return(filtered_cds)
})
 
# How many cells remain per sample? Does it match the expected cell number?
lapply(cds_list,function(x){dim(x)[2]})

lapply(filtered_cds_list,function(x){dim(x)[2]})

#Combine cds files
#cds<-combine_cds(filtered_cds_list, cell_names_unique = TRUE)
cds<-do.call(cbind,filtered_cds_list)

saveRDS(cds,"docall_combined_UMI500_cds.rds")
cds <- readRDS("docall_combined_UMI500_cds.rds")

colnames(fData(cds))<-c("gene_id")
rownames(cds)<-fData(cds)$gene_id
#colnames(pData(cds))[1]<-"barcode"
#rownames(pData(cds))<-paste(pData(cds)$barcode,pData(cds)$sample_id,sep=".")
colnames(cds)<-rownames(pData(cds))

# Summary stats
cds<-detect_genes(cds, min_expr = 0)
saveRDS(cds,"genes_detected_UMI500_cds.rds")
cds_test<-readRDS("genes_detected_UMI500_cds.rds")
rm(cds_test)
```

## Add gene-level annotation from BioMaRt (Get and add gene symbols from biomart)
Using the `gene_id` information in the featureData slot, we can fetch external annotations for each gene and merge them so we can get more meaningful gene information.  `biomaRt` is an interface in Bioconductor to get information associated with gene_ids.
```{r}
#Gene ID mappings using org.mm.eg are good but return "NA" for about half of ensembl id's, including many mitochondrial genes
#fData(cds)$gene_id_trimmed<-stringr::str_split_fixed(fData(cds)$gene_id,"\\.",2)[,1]
#symb <- mapIds(org.Mm.eg.db, keys=fData(cds)$gene_id_trimmed, keytype="ENSEMBL", column="SYMBOL")
#fData(cds)<-merge(fData(cds),as.data.frame(symb),by.x="gene_id_trimmed",by.y=0, all.x = TRUE,sort=FALSE)
#colnames(fData(cds))<-c("gene_id_trimmed","gene_id","num_cells_expressed","gene_short_name")

#Playing with BioMart, get gene names, use the right ensembl archive for genome used for alignment
listMarts()
listEnsemblArchives()
ensembl <- useEnsembl("ensembl", version = 100) 
#Used this to match with the previous version used by Dongeun
#datasets <- listDatasets(ensembl)
ensembl <- useDataset("mmusculus_gene_ensembl",mart=ensembl)

fData(cds)$gene_id_trimmed <- stringr::str_split_fixed(fData(cds)$gene_id,"\\.",2)[,1]

gene_description <- getBM(attributes = c("ensembl_gene_id_version","external_gene_name","description","chromosome_name"),
                   filters = "ensembl_gene_id", values = fData(cds)$gene_id_trimmed,  mart = ensembl)

#Prepare info as data frame for merging with sc object
rownames(gene_description) <- stringr::str_split_fixed(gene_description[,1],"\\.",2)[,1]

rownames(gene_description) <- gene_description[,1]

colnames(gene_description) <- c("gene_id","gene_short_name","description","chromosome_name")

fData(cds)
head(gene_description)

fData(cds) <- left_join(x=as.data.frame(fData(cds)),y=gene_description)

fData(cds)[fData(cds)$gene_id_trimmed == 'ENSMUSG00000000001',]
```

# metadata info
```{r}
batch<-c("2","3","4","5","6","7")
processing_date<-c("20190507","20190509","20200710","20200713","20200717","20230621")
age<-c("P30","P180","P360","P720")
sex<-c("M","F")
```

## Add cell-level annotations
```{r eval=make_cds}
metadata <- read.csv("2026_LncOL1_metadata_matn4.csv")
index <- metadata$sample_id
values <- c(7,7,7,2,3,4,5,6,6)
pData(cds)$batch <- values[as.factor(match(pData(cds)$sample_id,index))]

pData(cds)$processing_date <- processing_date[match(pData(cds)$batch,batch)]

index <- metadata$sample_id
age <- metadata$age
pData(cds)$age <- age[as.factor(match(pData(cds)$sample_id,index))]

sex <- metadata$sex
pData(cds)$sex <- sex[match(pData(cds)$sample_id,index)]

genotype <- metadata$genotype
pData(cds)$genotype <- genotype[match(pData(cds)$sample_id,index)]

pData(cds)
fData(cds)

save_monocle_objects(cds,"annotated_UMI500_cds.rds")
#cds_test <- load_monocle_objects("annotated_UMI500_cds.rds")
pData(cds_test)
fData(cds_test)
rm(cds_test)
```

# Choosing correct batches from Dongeun's dataset and our current dataset (Batch 7)
```{r}
table(pData(cds)$sample_id)
table(pData(cds)$batch)
cds <- cds[,pData(cds)$batch %in% c(2,3,5,7)]
table(pData(cds)$sample_id)
table(pData(cds)$batch)
```


# Supplementary Data 3- Plotting QC metrics before removing any cells
## Supplementary Data 3a- Number of genes expressed per cell
```{r}
pdf ("F:/orion_copies/kallisto_out_lncol1_matn4_sort/num_genes_p_cell.pdf", width = 6, height = 6)
hist(pData(cds)$num_genes_expressed,col="darkgreen",breaks=50,main="Number of genes expressed per cell", ylim = c(0,3500), xlim = c(0,10000))
dev.off()
```
## Supplementary Data 3b- Number of genes expressed per cell
```{r}
pdf ("F:/orion_copies/kallisto_out_lncol1_matn4_sort/num_cells_exp_gene.pdf", width = 6, height = 6)
hist(log10(fData(cds)$num_cells_expressed),col="red",breaks=50,main="Number of cells expressing a given gene", xlim = c(0,5) )
numCells<-10
abline(v=log10(numCells),lty="dashed")
dev.off()
```

## Supplementary Data 3c- Mitochondrial ratio
```{r}

pdf ("F:/orion_copies/kallisto_out_lncol1_matn4_sort/mito_ratio_0.05.pdf", width = 6, height = 6)

mito_genes<-fData(cds)$gene_short_name[grep("^mt-",fData(cds)$gene_short_name)]
mt_rows <- grepl("^mt-", fData(cds)$gene_short_name, ignore.case = TRUE)
sum(mt_rows) 
pData(cds)$mt_reads <- Matrix::colSums(exprs(cds)[mt_rows, , drop = FALSE])
pData(cds)$total_reads  <- Matrix::colSums(exprs(cds))
pData(cds)$mito_ratio <- pData(cds)$mt_reads/pData(cds)$total_reads
head(pData(cds)$mito_ratio )
hist(pData(cds)$mito_ratio,col="darkgreen",breaks=50,main="mito", ylim = c(0, 20000), xlim = c(0,1))

abline(v=0.05,lty="dashed")

dev.off()
```
## Supplementary Data 3d- Library saturation
```{r}

pdf ("F:/orion_copies/kallisto_out_lncol1_matn4_sort/library_saturation.pdf", width = 6, height = 6)

ggplot(as.data.frame(pData(cds)),aes(x=log10(total_reads),y=log10(num_genes_expressed))) +
  geom_bin2d(bins = 50) +
  scale_fill_continuous(type = "viridis") +

  ggtitle("Library Saturation")


dev.off()
```


# Remove genes with no or low detectable signal
```{r eval=make_cds}
numCells<-10
cds<-cds[fData(cds)$num_cells_expressed >= numCells,]
```

# Save cds file
```{r save_cds, eval=make_cds}
save_monocle_objects(cds,"annotated_UMI500_cds_filtered_corr")
```

## (Re)Load CDS (shortcut)
```{r reload_cds}
cds<-load_monocle_objects("annotated_UMI500_cds_filtered_corr")
```

## Pre-QC summary
Here are the per-condition cells/nuclei _*before*_ any QC controls (ie. raw yield from scRNA-Seq workflow).  
```{r}
kable(table(pData(cds)$age,pData(cds)$sex))
kable(table(pData(cds)$batch,pData(cds)$age))
kable(table(pData(cds)$age,pData(cds)$genotype))
```

# Moving to OneDrive for QC
```{r}
setwd("C:/Users/User/OneDrive - Johns Hopkins/Documents/R_Data/96Rik_w_Matn4/Re-analysis/for_revisions")
save_monocle_objects(cds,"annotated_UMI500_cds_filtered.rds")
```

## Load object
```{r}
#copied the following directly into the console
cds<-load_monocle_objects("annotated_UMI500_cds_filtered.rds")
pData(cds)
fData(cds)
```

### QC metrics
Next we can actually start to assess some of the important quality metrices for each gene and each cell.

#### Get Expressed Genes
```{r gene_qc_1}
#First, updating the UMI cutoff to the stricter cutoff based on the original knee plots
cds <- cds[,pData(cds)$Total_UMIs > 1000]

numCells<-10
cds<-detect_genes(cds)
expressed_genes <- row.names(subset(fData(cds),
    num_cells_expressed >= numCells))

length(expressed_genes)
```

Lets log transform and look again. This time, we'll add a threshold line showing our cutoff for 'expressed genes'.
```{r num_cells_expressing_log}
hist(log10(fData(cds)$num_cells_expressed),col="red",breaks=50,main="log10 Number of cells expressing a given gene")
abline(v=log10(numCells),lty="dashed")
```

We have now identified a total of `r length(expressed_genes)` that are detectably expressed in at least `r numCells` cells in our dataset.

#### Distribution of gene mean copies per cell
What is the average expression level (in mRNA Copies per cell) for each gene?
```{r gene_qc_2}
fData(cds)$mean_cpc<-Matrix::rowMeans(exprs(cds))
hist(log10(fData(cds[expressed_genes,])$mean_cpc),col="purple",breaks=50,main="Mean RNA copies per cell")

```

### Cell QC metrics

#### Distribution of detected genes across cells/age
How many genes are expressed in a given cell/age?
```{r cell_qc_1}
hist(pData(cds)$num_genes_expressed,col="darkgreen",breaks=50,main="Number of genes expressed per cell")

#Skipped
#p.nGenes<-ggplot(as.data.frame(pData(rik)),aes(x=log10(num_genes_expressed),fill=age)) +
#  geom_density() +
#  facet_grid(sample) +
#  scale_fill_brewer(palette="Set1") + 
#  monocle3:::monocle_theme_opts() +
#  ggtitle("Distribution of Log10 genes detected per cell")

#p.nGenes
```

#### Mt-genome proportion
A high proportion of mitochondrial genes may indicate a lower than ideal capture efficiency for a given cell. Here we identify the subset of mitochondrial genes and look at the proportion of reads mapping to 'mt-' genes vs genomic genes.
```{r cell_qc_mito}
mito_genes<-fData(cds)$gene_short_name[grep("^mt-",fData(cds)$gene_short_name)]
mt_rows <- grepl("^mt-", fData(cds)$gene_short_name, ignore.case = TRUE)
sum(mt_rows) 
pData(cds)$mt_reads <- Matrix::colSums(exprs(cds)[mt_rows, , drop = FALSE])
pData(cds)$total_reads  <- Matrix::colSums(exprs(cds))
pData(cds)$mito_ratio <- pData(cds)$mt_reads/pData(cds)$total_reads
head(pData(cds)$mito_ratio )
hist(pData(cds)$mito_ratio,col="darkgreen",breaks=50,main="mito")


save_monocle_objects(cds,"annotated_UMI1000_cds_filtered_mito_ratio.rds")
cds<-load_monocle_objects("annotated_UMI1000_cds_filtered_mito_ratio.rds")
```


# remove >50% mito-ratio
```{r}
#This is very high! I kept it so high to see if we could detect dying cells in our analysis- Later I moved this down to the standard 5%
cds <- cds[,!pData(cds)$mito_ratio > 0.50]
```

#### Total mRNAs per cell
```{r total_mRNAs}
pData(cds)$total_reads<-Matrix::colSums(exprs(cds))

cds <- cds[,pData(cds)$total_reads < 1e6]
upper_bound <- 10^(mean(log10(pData(cds)$total_reads)) +
                    2*sd(log10(pData(cds)$total_reads)))
lower_bound <- 10^(mean(log10(pData(cds)$total_reads)) -
                    2*sd(log10(pData(cds)$total_reads)))


p.UMIs<-ggplot(as.data.frame(pData(cds)),aes(x=log10(total_reads))) +
  geom_density() +

  scale_fill_brewer(palette="Set1") + 
  monocle3:::monocle_theme_opts() +
  ggtitle("Distribution of Log10 Total UMIs (mRNAs) recovered per cell") +
          geom_vline(xintercept = log10(lower_bound)) +
        geom_vline(xintercept = log10(upper_bound))

p.UMIs

```

#### Library Saturation curves

```{r}
p.sat<-ggplot(as.data.frame(pData(cds)),aes(x=log10(total_reads),y=log10(num_genes_expressed))) +
  geom_bin2d(bins = 50) +
  scale_fill_continuous(type = "viridis") +

  ggtitle("Library Saturation")

p.sat
```
#Identifying all non_OL in dataset
```{r Removing non-oligo lineage cells from the 96Rik dataset}
rownames(colData(cds)) <- colnames(cds)
cds <- estimate_size_factors(cds)
cds <- preprocess_cds(cds, num_dim = 100)
plot_pc_variance_explained(cds)
cds <- preprocess_cds(cds, num_dim = 26)
#See an inflection point around 26
cds <- reduce_dimension(cds)
plot_cells(cds)

plot_cells(cds,genes = c('Pdgfra','Matn4','Cspg4', "Olig2","Mbp", "Rbfox3","Gad1","Slc32a1","Syp", "Gfap", "Aqp4"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)
plot_cells(cds,genes = c('Olig2'), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)

colnames(pData(cds))[1]<-"barcode"
rownames(pData(cds))<-paste(pData(cds)$barcode,pData(cds)$sample_id,sep=".")
colnames(cds)<-rownames(pData(cds))
cds <- cluster_cells(cds, k= 45)


save_monocle_objects(cds,"20260204_cds_pre_batch_corr_UMI1000.rds")
#cds<-load_monocle_objects("20260204_cds_pre_batch_corr.rds")


plot_cells(cds, color_cells_by = "cluster", label_cell_groups = FALSE )

marker_test_res <- top_markers(cds, group_cells_by="cluster", 
                               reference_cells=1000)

gene_short_name <- fData(cds)$gene_short_name
gene_id <- fData(cds)$gene_id
marker_test_res$gene_short_name <- gene_short_name[match(marker_test_res$gene_id,gene_id)]


write.csv(marker_test_res, "20260204_marker_cluster_genes_contamination_UMI1000.csv")
save_monocle_objects(cds,"20260204_cds_pre_batch_corr_marker_genes_UMI1000.rds")

#1- OL
#2- OL
#3- OL
#4- OL
#5- OL
#6- OL
#7- OL
#8- microglia
#9- OL
#10- Neuron
#11- OL
#12- Endothelial
#13- OL
#14- Neuron
#15- Astrocyte



cl <- clusters(cds, reduction_method = "UMAP")
idx <- as.character(cl) %in% as.character(1:7,9,11,13)
colData(cds)$coarse_id <- NA_character_
colData(cds)$coarse_id[idx] <- "ol_lin"

idx <- as.character(cl) %in% as.character(12)
colData(cds)$coarse_id[idx] <- "endothelial"

idx <- as.character(cl) %in% as.character(10,14)
colData(cds)$coarse_id[idx] <- "neuron"

idx <- as.character(cl) %in% as.character(15)
colData(cds)$coarse_id[idx] <- "astrocyte"

idx <- as.character(cl) %in% as.character(8)
colData(cds)$coarse_id[idx] <- "microglia"

save_monocle_objects(cds,"20260207_cds_coarse_ids_UMI1000.rds")
#cds <- load_monocle_objects("20260207_cds_coarse_ids_UMI1000.rds")
```

## ED Fig. 3g- Pie plotting of the non-OL cells in the dataset
```{r Plot pie chart of the non-OL cells}
#grouping together all contamination for simplicity
library(ggplot2)

cl <- clusters(cds, reduction_method = "UMAP")
idx <- as.character(cl) %in% as.character(1:7,9,11,13)
colData(cds)$v_coarse_id <- NA_character_
colData(cds)$v_coarse_id[idx] <- "ol_lin"

idx <- as.character(cl) %in% as.character(12)
colData(cds)$v_coarse_id[idx] <- "non-ol"

idx <- as.character(cl) %in% as.character(10,14)
colData(cds)$v_coarse_id[idx] <- "non-ol"

idx <- as.character(cl) %in% as.character(15)
colData(cds)$v_coarse_id[idx] <- "non-ol"

idx <- as.character(cl) %in% as.character(8)
colData(cds)$v_coarse_id[idx] <- "non-ol"

#pie for lncol1
coarse_ids <- table(pData(cds)$v_coarse_id[pData(cds)$genotype == "lncol1_egfp"])

counts <- coarse_ids
pct <- 100 * counts / sum(counts)
pct
labs <- paste0(names(counts), " (", round(pct, 1), "%)")
cb <- c( "#CC79A7", "#009E73")
p <- (pie(counts,
labels = NA,
col = cb,
border = NA,
clockwise = TRUE,
init.angle = 90))


df <- data.frame(cat = names(coarse_ids), val = as.numeric(coarse_ids))
library(ggplot2)
p <- ggplot(df, aes(x = 1, y = val, fill = cat)) + geom_col(width = 1, color = NA) + coord_polar(theta = "y") + scale_fill_manual(values = cb) + theme_void()
ggsave("coarse_ids_donut_lncol1.png", plot = p, width = 6, height = 6, bg = "transparent")

#pie for matn4
coarse_ids <- table(pData(cds)$v_coarse_id[pData(cds)$genotype == "matn4_megfp"])

counts <- coarse_ids
pct <- 100 * counts / sum(counts)
pct
labs <- paste0(names(counts), " (", round(pct, 1), "%)")
cb <- c( "#CC79A7", "#009E73")
p <- (pie(counts,
labels = NA,
col = cb,
border = NA,
clockwise = TRUE,
init.angle = 90))


df <- data.frame(cat = names(coarse_ids), val = as.numeric(coarse_ids))
library(ggplot2)
p <- ggplot(df, aes(x = 1, y = val, fill = cat)) + geom_col(width = 1, color = NA) + coord_polar(theta = "y") + scale_fill_manual(values = cb) + theme_void()
ggsave("coarse_ids_donut_matn4.png", plot = p, width = 6, height = 6, bg = "transparent")

#saving a csv
coarse_ids <- table(pData(cds)$coarse_id[pData(cds)$genotype == "lncol1_egfp"])

write.csv(coarse_ids, "coarse_ids.csv")

```

#Selecting only OL lineage cells
```{r }
cds <- cds[,cds@clusters@listData[["UMAP"]][["clusters"]] %in% c(1:7,9,11,13)]

plot_cells(cds, color_cells_by ="cluster")
pData(cds)$batch <- as.character(pData(cds)$batch)
plot_cells(cds, color_cells_by ="batch")
plot_cells(cds, color_cells_by ="mito_ratio")

cds_low_mito <- cds[,!pData(cds)$mito_ratio > 0.05]


#confirming they're all expressing markers of the OL lineage
plot_cells(cds,genes = c('Olig2'), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)
plot_cells(cds_low_mito,genes = c('Olig2'), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)
plot_cells(cds_low_mito, color_cells_by ="mito_ratio")

save_monocle_objects(cds_low_mito, "20260207_cds_ol_lin_only_low_mito.rds")
```

#Testing Different Alignment strategies
##Align CDS according to batch only
```{r Aligning just based on batch which was converted to a character rather than a factor}
library(BiocParallel)
#cds_low_mito <- load_monocle_objects("20260207_cds_ol_lin_only_low_mito.rds")

cds_lo <- cds_low_mito
rm(cds_low_mito)
pData(cds_lo)$batch <- as.character(pData(cds_lo)$batch)
cds_lo <- align_cds(cds_lo, alignment_group = "batch")

cds_lo <- detect_genes(cds_lo)
cellCutoff<-10

expressed_genes <- row.names(subset(fData(cds_lo),
    num_cells_expressed >= cellCutoff))

expressed_genes <- intersect(row.names(rowData(cds_lo)), expressed_genes)

cds_lo <- reduce_dimension(cds_lo,
                        max_components=2,
                        preprocess_method = "Aligned",
                        reduction_method="UMAP",
                        umap.min_dist = 0.1,
                        umap.metric="cosine",
                        umap.n_neighbors= 15L,
                        umap.fast_sgd=FALSE,
                        cores=1
                        )

library(ggplot2)

plot_cells(cds_lo,genes = c('Pdgfra','Matn4','Cspg4', "Bmp4","Enpp6","9630013A20Rik", "Olig2","Mbp","Aspa"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)

plot_cells(cds_lo,color_cells_by="batch",cell_size=0.75, label_cell_groups=FALSE, alpha = 0.3)

plot_cells(cds_lo,color_cells_by="mito_ratio",cell_size=0.75, label_cell_groups=FALSE, alpha = 0.3)

save_monocle_objects(cds_lo, "20260207_align_batch_5mt")
```


```{r}
plot_cells(cds_lo,genes = c('Pdgfra','Matn4','Cspg4', "Bmp4","Enpp6","9630013A20Rik", "Olig2","Mbp","Aspa"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)

plot_cells(cds_lo,genes = c('Epcam'), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)

cds_lo <- cluster_cells(cds_lo, resolution = 1e-4)

plot_cells(cds_lo,color_cells_by = "cluster", norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)

marker_test_res <- top_markers(cds_lo, group_cells_by="cluster", 
                               reference_cells=1000)

gene_short_name <- fData(cds_lo)$gene_short_name
gene_id <- fData(cds_lo)$gene_id
marker_test_res$gene_short_name <- gene_short_name[match(marker_test_res$gene_id,gene_id)]

write.csv(marker_test_res, "20260207_marker_cluster_genes_lineage.csv")

plot_cells(cds_lo,color_cells_by = "cluster", norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE)  


cluster <- c(1,2,3,4,5,6,7,8,9,10,11,12,13)
identity<-c("6_myelin_forming_ol", #1
            "3_quiescent_OPC", #2
            "3_quiescent_OPC", #3
            "4_differentiating_OPC", #4
            "6_myelin_forming_ol", #5
            "6_myelin_forming_ol", #6
            "1_early_cycling_OPC", #7
            "7_mito_opc", #8
            "5_newly_formed_ol", #9
            "5_newly_formed_ol", #10
            "2_late_cycling_OPC", #11
            "9_mito_ol", #12
            "8_mito_diff" #13
            )

pData(cds_lo)$identity <- identity[match(cds_lo@clusters@listData[["UMAP"]][["clusters"]],cluster)]
pData(cds_lo)
plot_cells(cds_lo,color_cells_by = "identity", label_cell_groups = FALSE, alpha = 0.2)

save_monocle_objects(cds_lo, "20260207_cds_labeled.rds")
```

##Sorting out low quality cell groups
```{r}
cds_lo <- load_monocle_objects("20260204_cds_labeled.rds")

#two identities appear to have higher mitochondrial ratios- doing tests to determine their overall quality

#testing to see if any are significantly different groups based on mito_ratio
ggplot2::ggplot(as.data.frame(SummarizedExperiment::colData(cds_lo)), ggplot2::aes(x = identity, y = mito_ratio, fill = identity)) + ggplot2::geom_violin(scale = "width", trim = FALSE) + ggplot2::geom_boxplot(width = 0.1, outlier.shape = NA, color = "gray30") + ggplot2::theme_classic() + ggplot2::labs(x = "Identity", y = "Mitochondrial ratio") + ggplot2::guides(fill = "none")

df <- as.data.frame(SummarizedExperiment::colData(cds_lo))
df <- df[!is.na(df$mito_ratio) & !is.na(df$identity), ]
df$identity <- as.factor(df$identity)

kw <- kruskal.test(mito_ratio ~ identity, data = df)
print(kw)
pw <- pairwise.wilcox.test(df$mito_ratio, df$identity, p.adjust.method = "BH", exact = FALSE)

pw_df <- reshape2::melt(pw$p.value, na.rm = TRUE)
names(pw_df) <- c("group1", "group2", "p_adj")
sig_pairs <- subset(pw_df, p_adj < 0.05)
print(sig_pairs)

library(dplyr)
as.data.frame(SummarizedExperiment::colData(cds_lo)) %>%
group_by(identity) %>%
summarize(n = n(), mean_mito = mean(mito_ratio, na.rm = TRUE), .groups = "drop") %>%
arrange(desc(mean_mito))

#testing to see if any are significantly different groups based on Total_UMI
ggplot2::ggplot(as.data.frame(SummarizedExperiment::colData(cds_lo)), ggplot2::aes(x = identity, y = Total_UMIs, fill = identity)) + ggplot2::geom_violin(scale = "width", trim = FALSE) + ggplot2::geom_boxplot(width = 0.1, outlier.shape = NA, color = "gray30") + ggplot2::theme_classic() + ggplot2::labs(x = "Identity", y = "Total_UMIs") + ggplot2::guides(fill = "none")

df <- as.data.frame(SummarizedExperiment::colData(cds_lo))
df <- df[!is.na(df$Total_UMIs) & !is.na(df$identity), ]
df$identity <- as.factor(df$identity)

kw <- kruskal.test(Total_UMIs ~ identity, data = df)
print(kw)
pw <- pairwise.wilcox.test(df$Total_UMIs, df$identity, p.adjust.method = "BH", exact = FALSE)

pw_df <- reshape2::melt(pw$p.value, na.rm = TRUE)
names(pw_df) <- c("group1", "group2", "p_adj")
sig_pairs <- subset(pw_df, p_adj < 0.05)
print(sig_pairs)

as.data.frame(SummarizedExperiment::colData(cds_lo)) %>%
group_by(identity) %>%
summarize(n = n(), mean_umi = mean(Total_UMIs, na.rm = TRUE), .groups = "drop") %>%
arrange(desc(mean_umi))

#Two of the unusual clusters have the lowest mito_ratio and number of UMIs,  the mito_OL and mito_OPC, respectively, and they are significantly different from all other groups. If you are interested in either of these clusters, feel free to leave them in, as they could represent interesting death pathways in OPCs or OLs!

cds_lo <- cds_lo[, !(as.character(pData(cds_lo)$identity) %in% c("7_mito_opc","9_mito_ol"))]
plot_cells(cds_lo,color_cells_by = "identity", norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 0.75, label_cell_groups = FALSE) 

#updated to merge mito_diff with diff

cds_lo$identity[cds_lo$identity =="8_mito_diff"] <- "4_differentiating_OPC"


#save_monocle_objects(cds_lo, "20260207_cds_labeled_cleaned.rds")
#cds_lo <- load_monocle_objects("20260207_cds_labeled_cleaned.rds")
```


# Coverting to CellxGene Compatible Object
```{r}
#BiocManager::install("zellkonverter")
library(zellkonverter)

aws_batch <- cds_lo

aws_batch@rowRanges@elementMetadata@listData[["gene_short_name"]] <- make.unique(aws_batch@rowRanges@elementMetadata@listData[["gene_short_name"]])


assay(aws_batch, "logcounts") <- normalized_counts(aws_batch)

assays(aws_batch)

assay(aws_batch, "counts") <- NULL

assays(aws_batch)

reducedDims(aws_batch)
reducedDims(aws_batch)[["X_PCA"]] <- reducedDims(aws_batch)[["PCA"]]
reducedDims(aws_batch)[["X_UMAP"]] <- reducedDims(aws_batch)[["UMAP"]]

reducedDims(aws_batch)

colData(aws_batch) <- colData(aws_batch)[c("Size_Factor","age","genotype","batch","sex","mito_ratio","identity")]

rowData(aws_batch)
rowData(aws_batch) <- rowData(aws_batch)["gene_short_name"]
rowData(aws_batch)

#BiocManager::install("scRNAseq")
#library(scRNAseq)
#AnnDataDependencies(version = "0.7.6")
#writeH5AD(aws_batch, file = "./20261002_completelin_unique.h5ad",version = "0.7.6",
#                         X_name = "logcounts", 
#                         skip_assays = F)

library(SingleCellExperiment)
sce <- as(aws_batch, "SingleCellExperiment")

#writeH5AD(sce, file = "20261002_completelin_unique.h5ad")


```


#For text- Cells/genotype
```{r}

cds_lo <- load_monocle_objects("20260207_cds_labeled_cleaned.rds")
kable(table(pData(cds_lo)$age,pData(cds_lo)$genotype))

```


#Figure 2
##Figure 2 b
```{r Generating figures for revisions}
library(ggplot2)
#Saved as 2_genotype_both



plot_cells(cds_lo, color_cells_by = "genotype", label_cell_groups = FALSE, show_trajectory_graph = FALSE, cell_size = 1.5, alpha = 0.7) + ggplot2::scale_color_manual(values = c("#119C72","#55B3E8"))

#2_genotype_matn
cds_matn <- cds_lo[,(as.character(pData(cds_lo)$genotype) %in% c("matn4_megfp"))]
plot_cells(cds_matn, color_cells_by = "genotype", label_cell_groups = FALSE, show_trajectory_graph = FALSE, cell_size = 1.5, alpha = 0.7) + ggplot2::scale_color_manual(values = c("#55B3E8"))
rm(cds_matn)

#2_genotype_lnc
cds_lnc <- cds_lo[,(as.character(pData(cds_lo)$genotype) %in% c("lncol1_egfp"))]
plot_cells(cds_lnc, color_cells_by = "genotype", label_cell_groups = FALSE, show_trajectory_graph = FALSE, cell_size = 1.5, alpha = 0.7) + ggplot2::scale_color_manual(values = c("#119C72"))
rm(cds_lnc)
```
##Figure 2 d
```{r}
#2_identity_colored
plot_cells(cds_lo, color_cells_by = "identity", label_cell_groups = FALSE,  cell_size = 1) + ggplot2::scale_color_manual(values = c("#882255","#C17242","#DDCC77","#44AA99","#88CCEE","#332288"))

#For text- Top marker genes by cluster
marker_test_res <- top_markers(cds_lo, group_cells_by="identity", 
                               reference_cells=1000)

top_specific_markers <- marker_test_res %>%
                            filter(fraction_expressing >= 0.10) %>%
                            group_by(cell_group)
write.csv(top_specific_markers, "20260408_markers_by_id.csv")

save_monocle_objects(cds_lo, "20260207_cds_labeled_cleaned.rds")
```
##Extended data Fig. 3h- Pie chart identities
```{r}
#pie for lncol1
#cds_lo <- load_monocle_objects("20260207_cds_labeled_cleaned.rds")

coarse_ids <- table(pData(cds_lo)$identity)

counts <- coarse_ids
pct <- 100 * counts / sum(counts)
pct
labs <- paste0(names(counts), " (", round(pct, 1), "%)")
cb <- c( "#882255","#C17242","#DDCC77","#44AA99","#88CCEE","#332288")
p <- (pie(counts,
labels = NA,
col = cb,
border = NA,
clockwise = TRUE,
init.angle = 90))


df <- data.frame(cat = names(coarse_ids), val = as.numeric(coarse_ids))
library(ggplot2)
p <- ggplot(df, aes(x = 1, y = val, fill = cat)) + geom_col(width = 1, color = NA) + coord_polar(theta = "y") + scale_fill_manual(values = cb) + theme_void()
ggsave("identities_lncol1.png", plot = p, width = 6, height = 6, bg = "transparent")
```



## Figure 2 c
```{r}
marker_genes <- c("Mcm6","Ccnb2","Pdgfra","9630013A20Rik","Bcas1","Tcf7l2", "Mal")
plot_cells(cds_lo, genes = marker_genes, label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 5) 

```

##Figure 2 e 

```{r}
cds_lo <- cluster_cells(cds_lo)
cds_lo <- learn_graph(cds_lo, use_partition = FALSE)
cds_lo
cds_lo <- order_cells(cds_lo)
library(viridis)
library(ggplot2)
plot_cells(cds_lo, label_cell_groups = FALSE, alpha = 0.3, color_cells_by = "pseudotime", show_trajectory_graph = FALSE)+ scale_color_gradientn(colors = viridis(256, option = "E"))

save_monocle_objects(cds_lo, "20260204_cds_labeled_cleaned_pseudotime.rds")
```
##Fig 2 f
```{r}
cds_lo <- load_monocle_objects("20260204_cds_labeled_cleaned_pseudotime.rds")
  
pseudo_genes <- c("Aspa","Mbp","Bcas1","Cspg4","Top2a")
pseudo_cds <- cds_lo[rowData(cds_lo)$gene_short_name %in% pseudo_genes]
pseudo_cds <- order_cells(pseudo_cds)
plot_genes_in_pseudotime(pseudo_cds, color_cells_by = "identity", min_expr = 0.5, cell_size = 1) + ggplot2::scale_color_manual(values = c("#882255","#C17242","#DDCC77","#44AA99","#88CCEE","#332288"))
rm(pseudo_cds)
rm(pseudo_genes)

library(ggplot2)
library(viridis)
plot_cells(cds_lo, label_cell_groups = FALSE, alpha = 0.3, color_cells_by = "mito_ratio", show_trajectory_graph = FALSE)+ scale_color_gradientn(colors = viridis(256, option = "F"))
```

##Fig 2 g

```{r}
dead_genes <- c("Pdcd4","Sh3rf1","Pim3")
dead_cds <- cds_lo[rowData(cds_lo)$gene_short_name %in% dead_genes]
dead_cds <- order_cells(dead_cds)
plot_genes_in_pseudotime(dead_cds, color_cells_by = "identity", min_expr = 1.5, cell_size = 1) + ggplot2::scale_color_manual(values = c("#882255","#CC6677","#DDCC77","#44AA99","#88CCEE","#332288"))
rm(dead_cds)
rm(dead_genes)
```

##Fig 2 h
```{r}
dead_genes <- c("Casp3","Pdcd4","Sh3rf1","Pim3")
plot_cells(cds_lo, genes = dead_genes, label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 5) 

#save_monocle_objects(cds_lo, "20260205_cds_low_mito.rds")
#cds_lo <- load_monocle_objects("20260205_cds_low_mito.rds")
```

#Supplementary Figure

##Sup 3e- Batch effect
```{r}
plot_cells(cds_lo, color_cells_by = "batch", show_trajectory_graph = FALSE, label_cell_groups = FALSE, cell_size = 2, alpha = 0.2)+ ggplot2::scale_color_manual(values = c("#648FFF","#785EF0","#DC267F","#FE6100"))

library(dplyr)
cd <- colData(cds_lo)
idx <- order(cd$batch == "2")  
cds_lo_reorder <- cds_lo[, idx]

plot_cells(cds_lo_reorder,color_cells_by = "batch",show_trajectory_graph = FALSE,label_cell_groups = FALSE,cell_size = 2,alpha = 0.8) + ggplot2::scale_color_manual(values = c("#648FFF", "#D2D2D2", "#D2D2D2", "#D2D2D2"))

cd <- colData(cds_lo)
idx <- order(cd$batch == "3")  
cds_lo_reorder <- cds_lo[, idx]

plot_cells(cds_lo_reorder,color_cells_by = "batch",show_trajectory_graph = FALSE,label_cell_groups = FALSE,cell_size = 2,alpha = 0.8) + ggplot2::scale_color_manual(values = c("#D2D2D2", "#785EF0", "#D2D2D2", "#D2D2D2"))

idx <- order(cd$batch == "5")  
cds_lo_reorder <- cds_lo[, idx]

plot_cells(cds_lo_reorder,color_cells_by = "batch",show_trajectory_graph = FALSE,label_cell_groups = FALSE,cell_size = 2,alpha = 0.8) + ggplot2::scale_color_manual(values = c("#D2D2D2", "#D2D2D2", "#DC267F", "#D2D2D2"))


idx <- order(cd$batch == "7")  
cds_lo_reorder <- cds_lo[, idx]

plot_cells(cds_lo_reorder,color_cells_by = "batch",show_trajectory_graph = FALSE,label_cell_groups = FALSE,cell_size = 2,alpha = 0.8) + ggplot2::scale_color_manual(values = c("#D2D2D2","#D2D2D2","#D2D2D2","#FE6100"))
```

##Sup 2f- Plotting sex
```{r}
#plotting sex as a variable
cd <- colData(cds_lo)
idx <- order(cd$sex == "M")  
cds_lo_reorder <- cds_lo[, idx]

plot_cells(cds_lo_reorder,color_cells_by = "sex",show_trajectory_graph = FALSE,label_cell_groups = FALSE,cell_size = 2,alpha = 0.8) + ggplot2::scale_color_manual(values = c("#D2D2D2","#005AB5"))

idx <- order(cd$sex == "F")  
cds_lo_reorder <- cds_lo[, idx]

plot_cells(cds_lo_reorder,color_cells_by = "sex",show_trajectory_graph = FALSE,label_cell_groups = FALSE,cell_size = 2,alpha = 0.8) + ggplot2::scale_color_manual(values = c("#DC3220","#D2D2D2"))

rm(cds_lo_reorder)
rm(cd)
```


##Extended data Fig 4a- Pdcd4 graph
```{r}
marker_genes <- c("Pdcd4")
plot_cells(cds_lo, genes = marker_genes, label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 5, cell_size = 1.5) 
```
##Extended data Fig 4b- Highlight differentiating cell cluster
```{r}
plot_cells(cds_lo, color_cells_by = "identity", label_cell_groups = FALSE, show_trajectory_graph = FALSE, cell_size = 1.5) + ggplot2::scale_color_manual(values = c("#D2D2D2","#D2D2D2","#D2D2D2","#44AA99","#D2D2D2","#D2D2D2"))
```

## ED 4c- Differentiating cluster by genotype
```{r}
library(viridis)
cds_blob <- cds_lo[,(as.character(pData(cds_lo)$identity) %in% c("4_differentiating_OPC","8_mito_diff"))]

plot_cells(cds_blob, show_trajectory_graph = FALSE, cell_size = 2, color_cells_by = "mito_ratio")+ scale_color_gradientn(colors = viridis(500, option = "rocket"))

cd <- colData(cds_blob)
idx <- order(cd$genotype == "matn4_megfp")  
cds_lo_reorder <- cds_blob[, idx]
plot_cells(cds_blob, color_cells_by = "genotype", label_cell_groups = FALSE, show_trajectory_graph = FALSE, cell_size = 2, alpha = 1) + ggplot2::scale_color_manual(values = c("#D2D2D2","#55B3E8"))

cd <- colData(cds_blob)
idx <- order(cd$genotype == "lncol1_egfp")  
cds_lo_reorder <- cds_blob[, idx]
plot_cells(cds_lo_reorder, color_cells_by = "genotype", label_cell_groups = FALSE, show_trajectory_graph = FALSE, cell_size = 2, alpha = 1) + ggplot2::scale_color_manual(values = c("#119C72","#D2D2D2"))

cds_blob <- cluster_cells(cds_blob)
plot_cells(cds_blob, show_trajectory_graph = FALSE, cell_size = 1, color_cells_by = "cluster")
cds_blob <- learn_graph(cds_blob, use_partition = FALSE, learn_graph_control = list(prune_graph= FALSE))
plot_cells(cds_blob,
           color_cells_by = "batch",
           label_cell_groups = FALSE,
           label_groups_by_cluster=FALSE,
           label_leaves=TRUE,
           show_trajectory_graph = TRUE,
           label_branch_points=TRUE)
```

## ED 4d Adding gene expression plots
```{r}
plot_cells(cds_blob,genes = c("Cspg4"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 2, label_cell_groups = FALSE)

plot_cells(cds_blob,genes = c("Bcas1"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 2, label_cell_groups = FALSE)

plot_cells(cds_blob,genes = c("9630013A20Rik"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 2, label_cell_groups = FALSE)

plot_cells(cds_blob,genes = c("Mbp"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 2, label_cell_groups = FALSE)

plot_cells(cds_blob,genes = c("Mog"), norm_method = c("log", "size_only"), show_trajectory_graph=FALSE, min_expr = 5, cell_size = 2, label_cell_groups = FALSE)
```

##ED 4e-  Mito ratio plot
```{r}
plot_cells(cds_blob, color_cells_by = "mito_ratio", show_trajectory_graph = FALSE, cell_size = 2.5, alpha = 0.75)  + scale_color_viridis(option = "magma")
```

##ED 4f- Mito violin plot
```{r}

magma_baby <- c("#A3319F","#F89078")

library(ggplot2)
plot_cells(cds_blob, color_cells_by = "identity", label_cell_groups = FALSE, show_trajectory_graph = FALSE, cell_size= 2) + scale_color_manual(values =  magma_baby)

ggplot(as.data.frame(pData(cds_blob)), aes(x= identity, y=mito_ratio, fill = identity)) + geom_violin() + scale_fill_manual(values =  c("#A3319F","#F89078")) + theme_classic() 


with(as.data.frame(pData(cds_blob)), shapiro.test(pData(cds_blob)$mito_ratio[cds_blob$identity == "8_mito_diff"]))

res <- wilcox.test(mito_ratio ~ identity, data = as.data.frame(pData(cds_blob)),
                   exact = FALSE)
res
#p-value = 4.743e-14

```


##ED 4g Pseudotime trajectory plot
```{r}
only_grey <- c("#d3d3d3","#d3d3d3")
plot_cells(cds_blob, color_cells_by = "genotype", 
          show_trajectory_graph = TRUE, 
          label_cell_groups = FALSE, 
          cell_size = 1.5, 
          alpha = 1, 
          label_branch_points = FALSE,
          label_roots = FALSE,
          label_leaves = FALSE,
          trajectory_graph_color = "black",
          trajectory_graph_segment_size = 2
          ) + scale_colour_manual(values=only_grey)
```


##Spatial correlation in the blob for finding co-regulated genes
```{r Graph test for co-regulated genes}
pr_graph_test_res_blob <- graph_test(cds_blob, neighbor_graph="knn")
pr_deg_ids_blob <- row.names(subset(pr_graph_test_res_blob, morans_I > 0.01 & q_value < 0.05))

gene_module_df_blob <- find_gene_modules(cds_blob[pr_deg_ids_blob], resolution= 1e-3)

library(tidyverse)

gene_id_info <- as.data.frame(fData(cds_blob))
colnames(gene_id_info)[1] <- "id"
gene_module_df_annot <- merge(gene_module_df_blob, gene_id_info, by="id")

plot_cells(cds_blob, 
           genes = gene_module_df_blob %>% filter(module %in% c(1:5)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = TRUE)

plot_cells(cds_blob, 
           genes = gene_module_df_blob %>% filter(module %in% c(6:10)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)

plot_cells(cds_blob, 
           genes = gene_module_df_blob %>% filter(module %in% c(11:14)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = TRUE)

write.csv(gene_module_df_annot, "20260206_knn_graph_test_cds_blob.csv")
```

##Looking for variability along the trajectory graph
```{r Graph test across the pseudotime trajectory}

cds_blob_pseudotime_pr <- graph_test(cds_blob, neighbor_graph = "principal_graph")
blob_pr_deg_ids_pseudo <- row.names(subset(cds_blob_pseudotime_pr, q_value <0.05))

save_monocle_objects(cds_blob, "20260206_cds_blob_analysis.rds")

gene_module_df_blob_pseudo <- find_gene_modules(cds_blob[blob_pr_deg_ids_pseudo],resolution=c(10^seq(-6,1)))
unique(head(gene_module_df_blob_pseudo$module))                                            
write.csv(gene_module_df_blob_pseudo, "20260206_pseudo_graph_test_cds_blob.csv")

#51

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo %>% filter(module %in% c(1:10)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo %>% filter(module %in% c(11:20)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo %>% filter(module %in% c(21:30)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo %>% filter(module %in% c(31:40)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo %>% filter(module %in% c(41:51)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)

```

##GO Terms analysis for all cells on Off Ramp
```{r Spreadsheets of GO Terms modules off ramp}
#loop for all of the "off ramp" modules
library(gprofiler2)

gene_id_info <- as.data.frame(fData(cds_blob))
colnames(gene_id_info)[1] <- "id"
gene_module_df_blob_pseudo_annot <- merge(gene_module_df_blob_pseudo, gene_id_info, by="id")

mod_num_all <- 1:51

for (i in mod_num_all) {
module_sp_oc <- gene_module_df_blob_pseudo_annot %>% filter(module %in% i)
module_sp_gp <- gost(module_sp_oc$gene_short_name, organism="mmusculus", evcodes = TRUE)

mod_sp_gp_sv <- data.frame(module_sp_gp[["result"]][["term_name"]],module_sp_gp[["result"]][["intersection"]])
#remember to change name here
#write_csv(mod_sp_gp_sv, sprintf("20241031_surv_mod.%d._gp_oc.csv", i))
write.csv(mod_sp_gp_sv, sprintf("20260206_all_mod%d_gp_oc_od.csv", i))
rm(module_sp_gp)
rm(mod_sp_gp_sv)
}

#Notes on interesting modules identified and potential genes of interest

#2- DNA damage response
#Dyrk1b,Hltf,Ikbkg,Apc,Ccng1,Ppp4r3b,Tlk2,Susd6,Mlh3,Ercc6l2,Rad21,Prkdc,Ddb1,Huwe1,Smarcc2,Rbbp5,Dclre1c,Traf6,Fam168a,Herc2,Usp9x,Smarca5,Zbtb7a,Setd7,Ncoa6,Psme4,Cebpg,Upf1,Atxn7l3,Marf1,Bod1l

#2- cellular response to stress
#Sp1,Npepps,Chordc1,Dyrk1b,Hltf,Cdkn1b,Ikbkg,Apc,Rhoa,Ccng1,Ppp4r3b,Tlk2,Susd6,Atxn3,Mlh3,Ercc6l2,Ywhaz,Rad21,Prkdc,Dnajc18,Ddb1,Vldlr,P4hb,Huwe1,Smarcc2,Rbbp5,Dclre1c,Atf6,Traf6,Ube4b,Fam168a,Eif2ak1,Herc2,Usp9x,Eif2ak3,Smarca5,Neo1,Zbtb7a,Fnip1,Socs5,Setd7,Kdm6a,Opa1,Ncoa6,Nploc4,Cpeb2,Psme4,Ermp1,Tet1,Crebrf,Ywhag,Cebpg,Upf1,Atxn7l3,Marf1,Bod1l,Vkorc1l1,Rgma,Dnajb14,Ccdc47


#8- DNA Damage Response
#Fmr1,Hipk1,Sirt1,Atxn7,Epc1,Xiap,Ints3,Ep400,Xpc,Smg1,Atrx,Mbtps1,Arid2,Morc2a,Emsy,Bclaf1,Ascc3,Zmynd8,Pik3r1,Ppp4r3a,Tlk1,Setd2,Kmt5b,Trim39,Ercc6,Mdm4,Ep300,Usp47,Mbtd1,Taok2,Dot1l,Chd2

#9- Cellular response to stress
#Uba1,Dynll1,Rpa3,Atf6b,Actr2,Ywhae,Snw1,Eif2s1,Psmc6,Rnaseh2b,Smarcd1,Ube2j2,Cdc5l,Hspd1,Pdcl3,Ubxn4,Nucks1,Parp1,Psmd14,Akirin2,Ube2k,Coq7,Ppp4c,Morf4l2,Rnf126,Nsmce4a,Oard1,Sem1,Ubqln2,Xrcc1,Hmgb1,Ube2d3,Pigbos1


#26- Cellular response to stress
#Fer,Pnpt1,Nf1,Ddx5,Sel1l,Crebbp,Gsk3b,Dyrk1a,Wac,Smc3,Sec16a,Dnajc10,Rad23b,Brd7,Terf2,Sesn3,Smarca4,Fem1b,Fbxo22,Uba5,Vhl,Mtmr3,Supt16,Topors,Mapkap1,Morc3,Dhx9,Vcpip1,Foxo3,Taok3,Bmpr2,Epc2

#26- Apoptotic process
#Tax1bp1,Fbxo11,Crkl,Ctnnb1,Med1,Akap1,Nf1,Ddx5,Rabep1,Zmynd11,Gsk3b,Axin1,Smad4,Gnaq,Zdhhc3,Dnajc10,Ddx47,Smarca4,Fem1b,Vhl,Topors,Prkci,Mapkap1,Usp53,Foxo3,Adam10,Bmpr2,Epc2

#33- DNA damage response
#Baz1b,Cbx5,Top2b,Ppm1d,Smc6,Ercc4,Ube2v2,Usp7,Smchd1,Spire1,Skil,Sfpq,Casp2,Smarcad1,Rbbp6,Smarcc1,Atm,Rif1,Kdm1a,Slx4,Nsd2

#37- Positive regulation of cell death
#Dusp6,Dusp1,Grk2,Cflar,Stk17b,Tnfrsf1a,Btg1,Egr1,Mcl1,Ripk2,Lyn,Plagl2,Jun,Arrb2,Tnfaip8

#Cellular Response to Type II Interferon
#Ifngr1,Ifngr2,Jak1,Rab20,Sirpa,Rab11fip5
#Regulation of apoptotic process
#Zfp385a,Ier3,Mertk,Dusp6,Btg2,Dusp1,Grk2,Lrp5,Cflar,Stk17b,Nfkb1,Tnfrsf1a,Herpud1,Btg1,Egr1,Mcl1,Ripk2,Lyn,Plagl2,Jun,Arrb2,Tnfaip8

#Apoptosis
#Cflar,Nfkb1,Tnfrsf1a,Irf2,Mcl1,Jun

#38- Regulation of PCD


#39- Stress
#Uhrf1,Clgn,Paxip1,Msh6,Lnpk,Rdm1,Rnf168,Etaa1,Fbxo5,Myb,Pole2,Vrk1,Dek,Dhfr,Rangap1,Slc29a1,Polh,Bard1,Mcm6,Pcna,Smc2,Usp1,Rpa2,Xrcc2,Ung,Ezh2,Mcm7,Blm,Ipo5,Tipin,Topbp1,Smc4,Tdp2,Tmem97,Wdhd1,Smc1a,Spidr,H2ax,Rnaseh2a,Skp2,Hmgb2,Lig1,Cenps

#39- Apoptosis induced fragmentation
#apop_frag <- c("H1f4","Hmgb2",'H1f5')

#41- Stress
#Qrich1,Usp15,Cpeb4,Parg,Mapk8ip3,Alkbh8,Atf2,Ulk1,Rnf111,Ireb2,Cep63,Map2k4,Cbl,Otud4,Ubr5,Vps41,Rragb,Setx,Klhl15,Aff4,Brsk2,Kdm2a,Arid1b

#45- Viruses

#48- 
#regulation of programmed cell death
#apoptotic process
#regulation of apoptotic process

#51- Many of inflammation markers etc
```

##Reviewer comment- Middle of UMAP cells
```{r}
cds_blob <- load_monocle_objects("20260206_cds_blob_analysis.rds")

plot_cells(cds_blob, color_cells_by = "identity")

coarse_ids <- table(pData(cds_blob)$batch[pData(cds_blob)$identity == "8_mito_diff"])

counts <- coarse_ids
pct <- 100 * counts / sum(counts)
pct
labs <- paste0(names(counts), " (", round(pct, 1), "%)")
cb <- c("#648FFF","#785EF0","#DC267F","#FE6100")
p <- (pie(counts,
labels = NA,
col = cb,
border = NA,
clockwise = TRUE,
init.angle = 90))


df <- data.frame(cat = names(coarse_ids), val = as.numeric(coarse_ids))
library(ggplot2)
p <- ggplot(df, aes(x = 1, y = val, fill = cat)) + geom_col(width = 1, color = NA) + coord_polar(theta = "y") + scale_fill_manual(values = cb) + theme_void()
ggsave("mito_cluster_by_batch.png", plot = p, width = 6, height = 6, bg = "transparent")

#All cells

coarse_ids <- table(pData(cds_lo)$batch)

counts <- coarse_ids
pct <- 100 * counts / sum(counts)
pct
labs <- paste0(names(counts), " (", round(pct, 1), "%)")
cb <- c("#648FFF","#785EF0","#DC267F","#FE6100")
p <- (pie(counts,
labels = NA,
col = cb,
border = NA,
clockwise = TRUE,
init.angle = 90))


df <- data.frame(cat = names(coarse_ids), val = as.numeric(coarse_ids))
library(ggplot2)
p <- ggplot(df, aes(x = 1, y = val, fill = cat)) + geom_col(width = 1, color = NA) + coord_polar(theta = "y") + scale_fill_manual(values = cb) + theme_void()
ggsave("all_clusters_by_batch.png", plot = p, width = 6, height = 6, bg = "transparent")
```

##ED 4h- Death related modules
```{r Plot cells from interesting modules}
#gene_module_df_blob_pseudo <- read.csv("20260206_pseudo_graph_test_cds_blob.csv")

gene_id_info <- as.data.frame(fData(cds_blob))
colnames(gene_id_info)[1] <- "id"
gene_module_df_blob_pseudo_annot <- merge(gene_module_df_blob_pseudo, gene_id_info, by="id")

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo_annot %>% filter(module %in% c(12)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)
#Response to stress genes:
#Chmp1a,Ddx3x,Uba1,Ppard,Cul3,Letm1,Mtrex,Hsp90b1,Canx,Pnpt1,Ccnk,Tomm70a,Dyrk1a,Smarcd1,Cdc5l,Pja2,Pdpk1,Bag6,Grk2,Fundc1,Wdr45b,Bach1,Faf2,Nucks1,Vamp4,Ppp3ca,Map3k7,Ncbp1,Rad23b,Plaa,Pum1,Mtor,Rest,Bcl7a,Smarca4,Nedd4,Fem1b,Vhl,Mtmr3,Supt16,Fbxo45,Usp38,Mapkap1,Kat7,Foxk2,Pbrm1,Dhx9,Arf6,Eif4g1,Lrrc8d,Ythdf3,Zfp809,Chd4,Ipo7

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo %>% filter(module %in% c(36)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)
#Death receptor signalling
#Psen1,Ripk1,Mapk8,Nsmaf,Bag4,Arhgef11,Rnf31

plot_cells(cds_blob, 
           genes = gene_module_df_blob_pseudo %>% filter(module %in% c(38)),
           group_cells_by = "cluster", 
           color_cells_by = "cluster",
           show_trajectory_graph = FALSE)


#save_monocle_objects(cds_blob, "20260207_cds_blob")
#cds_blob <- load_monocle_objects("20260207_cds_blob")

```
## Matrisome analysis test
```{r}
#install.packages("dplyr")
#install.packages("data.table")
#install.packages("ggplot2")
#install.packages("ggalluvial")
#install.packages("crayon")
#devtools::install_github("izzilab/MatrisomeAnalyzeR")
library("MatrisomeAnalyzeR")
```
## Matrisome
```{r}
cds_lo <- load_monocle_objects("20260204_cds_labeled_cleaned_pseudotime.rds")

gene_short_name <- fData(cds_lo)$gene_short_name
gene_id <- fData(cds_lo)$gene_id

clusters <- c(0:7)
names(clusters) <- unique(colData(cds_lo)$identity)
for (i in clusters) {
  cells_i <- rownames(colData(cds_lo))[colData(cds_lo)$identity == names(clusters[i + 1])] 
  expr_i <- as.matrix(counts(cds_lo))[, cells_i, drop = FALSE]
  avg <- as.data.frame(rowMeans(expr_i, na.rm = TRUE)) %>% setNames("all")
  gns <- as.data.frame(rownames(subset(avg, all > 3)))
  colnames(gns) <- "Gene.Symbol"
  gns$Gene.Symbol <- gene_short_name[match(gns$Gene.Symbol,gene_id)]
  ann <- matriannotate(data = gns, gene.column = "Gene.Symbol", species = "mouse")
  nam <- paste("cluster", names(clusters[i + 1]), sep = "_")
  assign(nam, ann)
}

clusters_list <- list(
cluster_1_early_cycling_OPC = cluster_1_early_cycling_OPC,
cluster_2_late_cycling_OPC  = cluster_2_late_cycling_OPC,
cluster_3_quiescent_OPC     = cluster_3_quiescent_OPC,
cluster_4_differentiating_OPC = cluster_4_differentiating_OPC,
cluster_5_newly_formed_ol   = cluster_5_newly_formed_ol,
cluster_6_myelin_forming_ol = cluster_6_myelin_forming_ol
)
annot_col <- "Annotated Matrisome Category"
all_annot <- bind_rows(
lapply(names(clusters_list), function(nm) {
df <- clusters_list[[nm]]
if (!annot_col %in% colnames(df)) {
stop(sprintf("Column '%s' not found in %s", annot_col, nm))
}
df %>% mutate(Cluster = nm)
})
)
counts_per_cluster <- all_annot %>%
group_by(Cluster, .data[[annot_col]]) %>%
summarise(n = n(), .groups = "drop")

counts_per_cluster <-  counts_per_cluster[!(counts_per_cluster$`Annotated Matrisome Category` %in%c("Non-matrisome")),]
ggplot(counts_per_cluster, aes(x = Cluster, y = n, fill = .data[[annot_col]])) +
geom_bar(stat = "identity") +
labs(x = "Cluster", y = "Gene count", fill = "Matrisome category",
title = "Matrisome annotations per cluster") +
theme_classic() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


tab <- tidyr::pivot_wider(counts_per_cluster,
names_from = "Annotated Matrisome Category",
values_from = n,
values_fill = 0)
mat <- as.matrix(tab[ , setdiff(names(tab), "Cluster")])
rownames(mat) <- tab$Cluster
chisq.test(mat)

```

#Testing identified genes present at different stages
```{r}
plot_cells(cds_lo, genes = c("Col9a3"), label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 5, cell_size=2) 
plot_cells(cds_lo, genes = c("Serpine2"), label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 5, cell_size=2)
plot_cells(cds_lo, genes = c("Itih3"), label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 5, cell_size=2) plot_cells(cds_lo, genes = c("Plat"), label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 5, cell_size=2) 

```

##ED 3i-j- Violin plots over developmental trajectory
```{r}
#Using the old dataset for this
cds_old <- readRDS("C:/Users/User/OneDrive - Johns Hopkins/Documents/R_Data/96Rik_w_Matn4/Re-analysis/20231206_combined_matn4_96rik.rds")
cds_subset <- cds_old[rowData(cds_old)$gene_short_name %in% c("Pcdh17", "9630013A20Rik", "EGFP")]
cds_subset <- cds_subset[,grepl("lncOL1_EGFP", colData(cds_subset)$genotype)]
cds_subset <- cds_subset[,(as.character(pData(cds_subset)$identity) %in% c("4_late_diff_OPC","5_NFO","6_mature_OL"))]
plot_genes_violin(cds_subset, label_by_short_name = TRUE, group_cells_by = "identity")


gene_fits_identity <- fit_models(cds_subset, 
                             model_formula_str = "~num_genes_expressed+identity",
                             cores=2,
                             clean_model=TRUE,
                             verbose=TRUE
                             )

fit_coefs <- coefficient_table(gene_fits_identity)
id_terms <- fit_coefs %>% filter(term == "identity")
id_terms <- fit_coefs %>% filter(grepl("identity",term))

sig_id_terms<-id_terms %>% filter(q_value < 0.01 ) %>%
  dplyr::select(gene_id, gene_short_name, term, q_value, estimate, description)


write.csv(sig_id_terms,"C:/Users/User/OneDrive - Johns Hopkins/Documents/R_Data/96Rik_w_Matn4/Re-analysis/significant_id_terms.csv")


cds_subset <- cds_old[,grepl("lncOL1_EGFP", colData(cds_old)$genotype)]
cds_subset <- cds_subset[,(as.character(pData(cds_subset)$identity) %in% c("4_late_diff_OPC","5_NFO","6_mature_OL"))]
marker_genes <- c("Pcdh17","EGFP","9630013A20Rik","Opalin","Tcf7l2","Bcas1")
plot_cells(cds_subset, genes = marker_genes, label_cell_groups = FALSE, show_trajectory_graph = FALSE, min_expr = 2)

#Graph autocorrelation analysis
pr_graph_test_res <- graph_test(cds_subset)
pr_deg_ids <- row.names(subset(pr_graph_test_res, q_value < 0.05))
#All three vary significantly


gene_fits <- fit_models(cds_subset, model_formula_str = "~cluster")
fit_coefs <- coefficient_table(gene_fits)

```



#End of revisions code
```{r}
sessionInfo()

```
R version 4.5.2 (2025-10-31 ucrt)
Platform: x86_64-w64-mingw32/x64
Running under: Windows 11 x64 (build 26100)

Matrix products: default
  LAPACK version 3.12.1

locale:
[1] LC_COLLATE=English_United States.utf8 
[2] LC_CTYPE=English_United States.utf8   
[3] LC_MONETARY=English_United States.utf8
[4] LC_NUMERIC=C                          
[5] LC_TIME=English_United States.utf8    

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets 
[7] methods   base     

other attached packages:
 [1] Seurat_5.4.0                SeuratObject_5.3.0         
 [3] sp_2.2-0                    MatrisomeAnalyzeR_1.0.1    
 [5] ggalluvial_0.12.5           crayon_1.5.3               
 [7] data.table_1.18.2.1         dplyr_1.1.4                
 [9] ggplot2_4.0.2               monocle3_1.4.26            
[11] SingleCellExperiment_1.30.1 SummarizedExperiment_1.38.1
[13] GenomicRanges_1.60.0        GenomeInfoDb_1.44.3        
[15] IRanges_2.42.0              S4Vectors_0.46.0           
[17] MatrixGenerics_1.20.0       matrixStats_1.5.0          
[19] Biobase_2.68.0              BiocGenerics_0.54.1        
[21] generics_0.1.4             

loaded via a namespace (and not attached):
  [1] RcppAnnoy_0.0.22         later_1.4.5             
  [3] splines_4.5.2            BiocIO_1.18.0           
  [5] bitops_1.0-9             tibble_3.3.0            
  [7] polyclip_1.10-7          XML_3.99-0.20           
  [9] fastDummies_1.7.5        lifecycle_1.0.5         
 [11] Rdpack_2.6.5             globals_0.19.0          
 [13] processx_3.8.6           lattice_0.22-7          
 [15] ensembldb_2.32.0         MASS_7.3-65             
 [17] magrittr_2.0.3           plotly_4.12.0           
 [19] sass_0.4.10              rmarkdown_2.30          
 [21] jquerylib_0.1.4          yaml_2.3.10             
 [23] remotes_2.5.0            httpuv_1.6.16           
 [25] otel_0.2.0               sctransform_0.4.3       
 [27] spam_2.11-3              spatstat.sparse_3.1-0   
 [29] sessioninfo_1.2.3        pkgbuild_1.4.8          
 [31] reticulate_1.44.1        pbapply_1.7-4           
 [33] cowplot_1.2.0            DBI_1.2.3               
 [35] minqa_1.2.8              RColorBrewer_1.1-3      
 [37] lubridate_1.9.5          abind_1.4-8             
 [39] pkgload_1.5.0            Rtsne_0.17              
 [41] purrr_1.2.1              AnnotationFilter_1.32.0 
 [43] RCurl_1.98-1.17          GenomeInfoDbData_1.2.14 
 [45] ggrepel_0.9.6            irlba_2.3.7             
 [47] spatstat.utils_3.2-1     listenv_0.10.0          
 [49] goftest_1.2-3            RSpectra_0.16-2         
 [51] spatstat.random_3.4-4    fitdistrplus_1.2-6      
 [53] parallelly_1.45.1        codetools_0.2-20        
 [55] DelayedArray_0.34.1      tidyselect_1.2.1        
 [57] UCSC.utils_1.4.0         farver_2.1.2            
 [59] lme4_1.1-37              viridis_0.6.5           
 [61] spatstat.explore_3.7-0   GenomicAlignments_1.44.0
 [63] jsonlite_2.0.0           ellipsis_0.3.2          
 [65] progressr_0.18.0         ggridges_0.5.7          
 [67] survival_3.8-3           tools_4.5.2             
 [69] ica_1.0-3                Rcpp_1.1.0              
 [71] glue_1.8.0               gridExtra_2.3           
 [73] SparseArray_1.8.1        xfun_0.52               
 [75] usethis_3.2.1            withr_3.0.2             
 [77] fastmap_1.2.0            boot_1.3-32             
 [79] callr_3.7.6              digest_0.6.37           
 [81] mime_0.13                timechange_0.4.0        
 [83] R6_2.6.1                 scattermore_1.2         
 [85] tensor_1.5.1             spatstat.data_3.1-9     
 [87] RSQLite_2.4.5            tidyr_1.3.2             
 [89] rtracklayer_1.68.0       htmlwidgets_1.6.4       
 [91] httr_1.4.7               S4Arrays_1.8.1          
 [93] uwot_0.2.4               pkgconfig_2.0.3         
 [95] gtable_0.3.6             rsconnect_1.7.0         
 [97] blob_1.3.0               lmtest_0.9-40           
 [99] S7_0.2.1                 XVector_0.48.0          
[101] htmltools_0.5.8.1        dotCall64_1.2           
[103] plyranges_1.28.0         ProtGenerics_1.40.0     
[105] scales_1.4.0             png_0.1-8               
[107] spatstat.univar_3.1-6    reformulas_0.4.4        
[109] knitr_1.51               rstudioapi_0.18.0       
[111] reshape2_1.4.5           rjson_0.2.23            
[113] nlme_3.1-168             curl_7.0.0              
[115] nloptr_2.2.1             zoo_1.8-15              
[117] cachem_1.1.0             stringr_1.6.0           
[119] KernSmooth_2.23-26       miniUI_0.1.2            
[121] parallel_4.5.2           AnnotationDbi_1.70.0    
[123] restfulr_0.0.16          desc_1.4.3              
[125] pillar_1.11.1            grid_4.5.2              
[127] vctrs_0.6.5              RANN_2.6.2              
[129] promises_1.5.0           xtable_1.8-4            
[131] cluster_2.1.8.1          evaluate_1.0.5          
[133] GenomicFeatures_1.60.0   cli_3.6.5               
[135] compiler_4.5.2           Rsamtools_2.24.1        
[137] rlang_1.1.6              future.apply_1.20.1     
[139] labeling_0.4.3           ps_1.9.1                
[141] plyr_1.8.9               fs_1.6.6                
[143] stringi_1.8.7            deldir_2.0-4            
[145] viridisLite_0.4.3        BiocParallel_1.42.2     
[147] assertthat_0.2.1         Biostrings_2.76.0       
[149] lazyeval_0.2.2           spatstat.geom_3.7-0     
[151] devtools_2.4.6           Matrix_1.7-4            
[153] RcppHNSW_0.6.0           patchwork_1.3.2         
[155] bit64_4.6.0-1            future_1.69.0           
[157] shiny_1.12.1             KEGGREST_1.48.1         
[159] rbibutils_2.4.1          ROCR_1.0-12             
[161] igraph_2.1.4             memoise_2.0.1           
[163] bslib_0.10.0             bit_4.6.0  


